### 2023-09-22 paper/supplement.tex
finished adding meta analysis to main text, also added meta analysis tables to supplemental materials

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage[sort]{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
% \graphicspath{ {./figures-low/} }
\graphicspath{ {./figures-default/} }
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{authblk}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{multirow}
\usepackage[online,flushleft]{threeparttable}
\usepackage{booktabs}

\doublespacing
\linenumbers

\title{\textit{Supplementary materials for:} The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events}
\author[1]{Xinming Xu}
\author[2]{Ziyan Zhu}
\author[1, $\star$]{Jeremy R. Manning}
\affil[1]{Dartmouth College, Hanover, NH, USA}
\affil[2]{Peking University, Beijing, China}
\affil[$\star$]{Address correspondence to jeremy.r.manning@dartmouth.edu}

\begin{document}
\maketitle

\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

\begin{table}[]
\caption{\textbf{Stimuli description.}}
\begin{center}
\resizebox{\textwidth}{!}{
\begin{threeparttable}
\begin{tabular}{rrrrlll}
\toprule
\multirow{2}{*}{Storyline} &
  \multirow{2}{*}{Segment} &
  \multirow{2}{*}{Episode} &
  \multirow{2}{*}{Duration (s)} &
  \multirow{2}{*}{Main characters} &
  \multicolumn{2}{l}{Number of events\tnote{a}} \\ \cmidrule(l){6-7} 
  &    &   &     &                                & Onscreen      & Offscreen\tnote{b} \\ \midrule
1 & 1  & 1 & 105 & Beth, Sheila, Rob, Leo         & 4             & 7         \\
1 & 2  & 1 & 172 & Beth, Sheila, Rob, Leo         & 7 {[}1{]}     & 3         \\
1 & 3  & 1 & 70  & Beth, Sheila, One neighbor     & 6             & 2         \\
1 & 4  & 1 & 135 & Beth, Sheila                   & 8 {[}1{]}     & 1         \\
1 & 5  & 1 & 90  & Beth, Rob, April               & 8             & 4         \\
1 & 6  & 1 & 188 & Beth, Rob                      & 6 {[}2{]}     & 1         \\
1 & 7  & 1 & 91  & Beth, Sheila                   & 5 {[}1{]}     & 2         \\
1 & 8  & 1 & 142 & Beth, April                    & 4 (1) {[}2{]} & 3 (1)     \\
1 & 9  & 2 & 134 & Beth, April                    & 4 {[}2{]}     & 1         \\
1 & 10 & 2 & 58  & Beth                           & 3 {[}1{]}     & 6 {[}1{]} \\
1 & 11 & 2 & 159 & Beth, Rob                      & 9             &           \\
2 & 1  & 1 & 70  & Simone, Karl                   & 3 {[}1{]}     & 6         \\
2 & 2  & 1 & 119 & Simone, Karl, Naomi, Tommy     & 4 {[}1{]}     & 2         \\
2 & 3  & 1 & 58  & Simone, Karl, Tommy            & 4             & 2         \\
2 & 4  & 1 & 102 & Simone, Karl                   & 4             & 6         \\
2 & 5  & 1 & 134 & Simone, Karl, Tommy            & 6 (3) {[}1{]} & 3         \\
2 & 6  & 1 & 81  & Simone, Karl, Neighbors, Wanda & 4 (1)         & 6         \\
2 & 7  & 1 & 194 & Simone, Tommy                  & 8             & 3         \\
2 & 8  & 2 & 101 & Simone, Karl                   & 3 (1)         & 1         \\
2 & 9  & 2 & 86  & Simone, Tommy                  & 4 {[}1{]}     & 3         \\
2 & 10 & 2 & 232 & Simone, Karl                   & 6 {[}1{]}     & 3         \\
2 & 11 & 2 & 189 & Simone, Naomi, Tommy           & 7 (1) {[}1{]} &           \\ \bottomrule
\end{tabular}
\begin{tablenotes}
\item[a] Number of partial events (see methods) shown in parentheses; number of summary events shown in square brackets.
\item[b] Offscreen events happened between the current segment and the next segment.
\end{tablenotes}
\end{threeparttable}
}
\end{center}
\end{table}

\begin{table}
    \footnotesize
    \caption{List of datasets analyzed in our meta-analysis (see Fig. 8).}
    \label{tab:datasets}
    \begin{tabular}{p{0.6in}|l|p{2in}|l|p{0.5in}|p{0.5in}|p{0.5in}}
        \toprule
        \textbf{Dataset} & \textbf{Short name} & \textbf{Description} & \textbf{Category} & \textbf{Number of observations} & \textbf{Ob\-ser\-va\-tion type} & \textbf{Number of words} \\
        \midrule
        Internet Movie Script Database & IMSDb & A collection of transcripts from roughly 1000 popular movies. & Film & 1091 & Transcript & 26023348 \\
        \midrule
        Movie Dialogues Dataset & Movies & A large collection of fictional conversations extracted from raw movie scripts. & Film & 304713 & Utterance & 3209921 \\
        \midrule
        Switchboard Dialog Act Corpus & Switchboard & A collection of five-minute telephone conversations between two participants, annotated with speech act tags. & Speech & 122646 & Utterance & 2052779 \\
        \midrule
        Supreme Court Corpus & SCOTUS & A collection of cases from the U.S. Supreme Court, along with transcripts of oral arguments. & Speech & 1700789 & Utterance & 71889094 \\
        \midrule
        Tennis Interviews & Tennis & Transcripts for tennis singles post-match press conferences for major tournaments between 2007 to 2015. & Speech & 163948 & Utterance & 7043118 \\
        \midrule
        Persuasion for Good Corpus & PfG & A collection of online conversations generated by Amazon Mechanical Turk workers, where one participant (the persuader) tries to convince the other (the persuadee) to donate to a charity. & Speech & 20932 & Utterance & 351759 \\
        \midrule
        Intelligence Squared Debates Corpus & IQ2 & This dataset contains transcripts of debates held as part of Intelligence Squared Debates. & Speech & 26562 & Utterance & 1898509 \\
        \midrule
        Group Affect and Performance Corpus & GAP & Group members completed a Winter Survival Task (WST), a group decision-making exercise where participants must rank 15 items according to their importance in a hypothetical plane crash scenario. Participants first rank the items individually. Then, each group was given a maximum of 15 minutes to complete the WST. The group’s conversations and deliberations during this task were recorded as conversations in this dataset. & Speech & 8009 & Utterance & 45989 \\
        \midrule
        The Chair & Chair & Scraped transcripts from The Chair, Season 1. & Television & 6 & Transcript & 19197 \\
        \midrule
        Friends Corpus & Friends & A collection of all the conversations that occurred over 10 seasons of Friends, a popular American TV sitcom that ran in the 1990s. & Television & 67373 & Utterance & 622894 \\
        \midrule
        Gutenberg Dialogue Dataset & Gutenberg & Dialogues extracted from the Project Gutenberg collection. & Text & 14773741 & Utterance & 327519461 \\
        \midrule
        Reddit Corpus & Reddit & A collection of Corpuses of Reddit data built from Pushshift.io Reddit Corpus. Each Corpus contains posts and comments from an individual subreddit from its inception until Oct 2018. & Text & 74468 & Utterance & 3080662 \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{table}
    \footnotesize
    \centering

    \caption{Part of speech pattern templates used to identify verb tenses in
    our meta analysis. Part of speech tags are represented between angled
    brackets, and pipes ($|$) denote instances where \textit{any} of the two or
    more parts of speech are considered a match when they appear at the given
    position in the indicated sequence. The plus signs ($+$) denote that one or
    more repetitions of the given sequence within a single sentence are still
    counted as a ``single'' instance of the given tense. Part of speech tags
    are defined in Table~\ref{tab:pos_tags}.}

    \begin{tabular}{r|l}
    \toprule
      \textbf{Tense} & \textbf{Pattern template} \\
      \midrule
      conditional continuous & $<$MD$>$$<$BE$>$$<$VBG$|$HVG$|$BEG$>$$+$ \\
      \midrule
      conditional continuous passive & $<$MD$>$$<$BE$>$$<$BEG$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      conditional indefinite & $<$MD$>$$<$BE$|$DO$|$VB$|$HV$>$$+$ \\
      \midrule
      conditional indefinite passive & $<$MD$>$$<$BE$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      conditional perfect & $<$MD$>$$<$HV$>$$<$HVN$|$BEN$|$VBN$|$VBD$>$$+$ \\
      \midrule
      conditional perfect continuous & $<$MD$>$$<$HV$>$$<$BEN$>$$<$VBG$|$HVG$|$BEG$>$$+$ \\
      \midrule
      conditional perfect passive & $<$MD$>$$<$HV$>$$<$BEN$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      future continuous & $<$MDF$>$$<$BE$>$$<$VBG$|$HVG$|$BEG$>$$+$ \\
      \midrule
      future continuous passive & $<$MDF$>$$<$BE$>$$<$BEG$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      future indefinite & $<$MDF$>$$<$BE$|$DO$|$VB$|$HV$>$$+$ \\
      \midrule
      future indefinite passive & $<$MDF$>$$<$BE$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      future perfect & $<$MDF$>$$<$HV$>$$<$HVN$|$BEN$|$VBN$|$VBD$>$$+$ \\
      \midrule
      future perfect continuous & $<$MDF$>$$<$HV$>$$<$BEN$>$$<$VBG$|$HVG$|$BEG$>$$+$ \\
      \midrule
      future perfect continuous passive & $<$MDF$>$$<$HV$>$$<$BEN$>$$<$BEG$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      future perfect passive & $<$MDF$>$$<$HV$>$$<$BEN$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      infinitive & $<$TO$>$$<$BE$|$HV$|$VB$>$$+$ \\
      \midrule
      past continuous & $<$BED$|$BEDZ$>$$<$VBG$|$HVG$|$BEG$>$$+$ \\
      \midrule
      past continuous passive & $<$BED$|$BEDZ$>$$<$BEG$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      past indefinite & $<$DOD$>$$<$VB$|$HV$|$DO$>$$|$$<$BEDZ$|$BED$|$HVD$|$VBN$|$VBD$>$$+$ \\
      \midrule
      past indefinite passive & $<$BED$|$BEDZ$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      past perfect & $<$HVD$>$$<$BEN$|$VBN$|$HVD$|$HVN$>$$+$ \\
      \midrule
      past perfect continuous & $<$HVD$>$$<$BEN$>$$<$HVG$|$BEG$|$VBG$>$$+$ \\
      \midrule
      past perfect continuous passive & $<$HVD$>$$<$BEN$>$$<$BEG$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      past perfect passive & $<$HVD$>$$<$BEN$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      present continuous & $<$BEM$|$BER$|$BEZ$>$$<$BEG$|$VBG$|$HVG$>$$+$ \\
      \midrule
      present continuous passive & $<$BEM$|$BER$|$BEZ$>$$<$BEG$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      present indefinite & $<$DO$|$DOZ$>$$<$DO$|$HV$|$VB$>$$+$$|$$<$DO$|$HV$|$VB$|$BEZ$|$DOZ$|$BER$|$HVZ$|$BEM$|$VBZ$>$$+$ \\
      \midrule
      present indefinite passive & $<$BEM$|$BER$|$BEZ$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      present perfect & $<$HV$|$HVZ$>$$<$BEN$|$HVD$|$VBN$|$VBD$>$$+$ \\
      \midrule
      present perfect continuous & $<$HV$|$HVZ$>$$<$BEN$>$$<$VBG$|$BEG$|$HVG$>$$+$ \\
      \midrule
      present perfect continuous passive & $<$HV$|$HVZ$>$$<$BEN$>$$<$BEG$>$$<$VBN$|$VBD$>$$+$ \\
      \midrule
      present perfect passive & $<$HV$|$HVZ$>$$<$BEN$>$$<$VBN$|$VBD$>$$+$ \\
      \bottomrule
    \end{tabular}
    \label{tab:regexp_templates}
\end{table}

\begin{table}[h!]
    \centering
    \caption{Part of speech tag definitions.}
    \begin{tabular}{r|l}
        \toprule
        \textbf{Tag} & \textbf{Part of speech} \\
        \midrule
        BE & Be \\
        \midrule
        BEG & Related to beginning or gerundive form \\
        \midrule
        BEM & Am (first person singular present of BE) \\
        \midrule
        BEN & Been (past participle of BE) \\
        \midrule
        BER & Are (second person singular and all plural present of BE) \\
        \midrule
        BEZ & Is (third person singular present of BE) \\
        \midrule
        BED & Were (past plural of BE) \\
        \midrule
        BEDZ & Was (past singular of BE) \\
        \midrule
        DO & Do \\
        \midrule
        DOD & Did (past tense of DO) \\
        \midrule
        DOZ & Do, 3rd person singular present \\
        \midrule
        HV & Have (archaic form for historical texts) \\
        \midrule
        HVD & Have, past tense (archaic) \\
        \midrule
        HVG & Have, gerund or present participle (archaic) \\
        \midrule
        HVN & Have, past participle (archaic) \\
        \midrule
        HVZ & Have, 3rd person singular present (archaic) \\
        \midrule
        MD & Modal \\
        \midrule
        MDF & Modal, future tense \\
        \midrule
        TO & To \\
        \midrule
        VB & Verb, base form \\
        \midrule
        VBD & Verb, past tense \\
        \midrule
        VBG & Verb, gerund or present participle \\
        \midrule
        VBN & Verb, past participle \\
        \midrule
        VBP & Verb, non-3rd person singular present \\
        \midrule
        VBZ & Verb, 3rd person singular present \\
        \bottomrule
    \end{tabular}
    \label{tab:pos_tags}
  \end{table}

  \begin{table}[h!]
    \centering
    \caption{Keywords and phrases used to identify references to the past.}
    \begin{tabular}{l l l l l}
    \toprule
    ago & elapsed & last month & olden days & thus far \\
    \midrule
    already & expired & last night & once & to date \\
    \midrule
    antiquity & final & last quarter & once upon a time & up to now \\
    \midrule
    back when & formerly & last season & previously & used to \\
    \midrule
    before & had & last semester & recently & used to be \\
    \midrule
    bygone & heretofore & last time & said & was \\
    \midrule
    ceased & historically & last week & since & were \\
    \midrule
    concluded & hitherto & last year & so far & wrote \\
    \midrule
    did & in the past & long ago & terminated & yesterday \\
    \midrule
    earlier & in those days & made & then & yesteryear \\
    \bottomrule
    \end{tabular}
    \label{tab:keywords_past}
  \end{table}

  \begin{table}[h!]
    \centering
    \caption{Keywords and phrases used to identify references to the future.}
    \begin{tabular}{l l l l l}
        \toprule
        after & going to & later & next time & shortly \\
        \midrule
        anticipated & hereafter & later on & next week & some day \\
        \midrule
        can & imminently & looming & next year & soon \\
        \midrule
        could & impending & may & on the horizon & subsequent \\
        \midrule
        down the line & in the cards & might & plan to & succeeding \\
        \midrule
        eventual & in the future & next month & predicted & to be \\
        \midrule
        eventually & in the works & next quarter & prospective & tomorrow \\
        \midrule
        forthcoming & in time & next season & scheduled to & upcoming \\
        \midrule
        futuristic & intend to & next semester & shall & will \\
        \bottomrule
    \end{tabular}
    \label{tab:keywords_future}
  \end{table}

  \newpage

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp1.pdf}
    \caption{\textbf{Mean proportion of target events hit (A), precision (B), and convergence (C) as a function of number of segments watched in each storyline, in participants' uncued and character-cued retrodictions and predictions.} Grey lines represent the least squares fits.}
    \label{fig:supp1}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp2.pdf}
    \caption{\textbf{Mean number of events hit for each event type, in participants' uncued and character-cued retrodictions and predictions, averaged across just-watched segments.}}
    \label{fig:supp2}
\end{figure}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{supp3.pdf}
    \caption{\textbf{(A)} References in each storyline. \textbf{(B)} Distribution of reference lags.}
    \label{fig:supp3}
\end{figure}

\end{document}