### 2023-09-23 paper/main.tex
finished a complete round of edits (main text + supplement)

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
% \graphicspath{ {./figures-low/} }
% \DeclareGraphicsExtensions{.png,.pdf}
\graphicspath{ {./figures-default/} }
\usepackage{newfloat}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}
\usepackage{placeins}
\usepackage{authblk}
\usepackage{textcomp}
\usepackage{listings}
\bibliographystyle{apalike}

\doublespacing
\linenumbers

\title{\Large The psychological arrow of time drives temporal asymmetries in inferring unobserved past and future events}
\author[1]{Xinming Xu}
\author[2]{Ziyan Zhu}
\author[3]{Xueyao Zheng}
\author[1, $\star$]{Jeremy R. Manning}
\affil[1]{Dartmouth College, Hanover, NH, USA}
\affil[2]{Peking University, Beijing, China}
\affil[3]{Beijing Normal University, Beijing, China}
\affil[$\star$]{Address correspondence to jeremy.r.manning@dartmouth.edu}

% tables
\newcommand{\stimDescription}{S1}
\newcommand{\regExpTable}{S2}
\newcommand{\posTags}{S3}
\newcommand{\pastKeys}{S4}
\newcommand{\futureKeys}{S5}
\newcommand{\metaAnalysisDatasets}{S6}

% figures
\newcommand{\segments}{S1}
\newcommand{\events}{S2}
\newcommand{\references}{S3}
\newcommand{\MethodsReplExp}{S4}
\newcommand{\targetAsymmetries}{S5}
\newcommand{\hitRates}{S6}
\newcommand{\characterRefs}{S7}
\newcommand{\refAdjacent}{S8}
\newcommand{\referringReferenced}{S9}

\begin{document}

\maketitle

\begin{abstract} {\footnotesize{ How much can we infer about the past and
future, given our knowledge of the present? Unlike temporally symmetric
inferences about simple sequences, inferences about our own lives are
asymmetric: we are better able to infer the past than the future, since we
remember our past but not our future (i.e., the psychological arrow of time).
What happens when both the past and future are unobserved, as when we make
inferences about \textit{other} people's lives? We had participants in two
experiments view segments of two character-driven television dramas. They wrote
out what would happen just before or after each just-watched segment.
Participants were better at inferring past (versus future) events. This
asymmetry was driven by participants’ reliance on characters’ conversational
references in the narrative, which tended to favor the past. We also carried
out a meta analysis to estimate the prevalence of these asymmetries in hundreds
of millions of dialogues from television shows, popular movies, novels, and
written and spoken natural conversations. We found that, on average, references
to the past are roughly 1.5--2 times more prevalent in human conversations than
references to the future. Our work reveals a temporal asymmetry in how
observations of other people’s behaviors can inform us about the past and
future.

\textbf{Keywords: arrow of time, prediction, retrodiction, narrative, conversation}}}

\end{abstract}

\section*{Introduction}

What we experience in the current moment tells us about \textit{now}-- but what
does it tell us about the past or future? And does the current moment tell us,
as human observers, \textit{more} about the past or about the future? One way
of examining these questions is to consider highly simplified scenarios that
are artificially constructed in the laboratory \citep[e.g., ][]{MaheEtal22}. At
one extreme, for deterministic sequences with \textit{known} rules, knowing the
current state provides the observer with sufficient information to exactly
reconstruct the entire past and future history of the stimulus. At another
extreme, for purely random sequences, observing the current state provides no
information about the past \textit{or} future.

Sequences generated by stochastic processes fall somewhere between these two
extremes. For Markov processes, where each state is solely dependent on the
immediately preceding state, Shannon entropy may be used to quantify the
uncertainty of the past and future states, given the present state.
\cite{Cove94} showed that, for any stationary process (i.e., processes in
equilibrium), Markov or otherwise, the present state provides equal information
(i.e., mutual information) about past and future states~\citep[also
see][]{BialEtal01, ElliEtal09}. Further, there is some evidence that humans are
similarly adept at inferring the most likely previous and next items in
sequences governed by stochastic Markov processes~\citep{JonePash07}.

Deterministic, random, and probabilistic sequences (in equilibrium) are all
symmetric: the present state of these sequences is equally informative about
past versus future states. In contrast, our subjective experience in everyday
life is that we know more about our own past than our
future~\citep[e.g.,][]{Horw87}. We have memories of our past that we carry with
us into the present moment, but we do not have memories of our
yet-to-be-experienced future. This temporal asymmetry imposes an ``arrow of
time'' on our subjective experience, known as the \textit{psychological arrow
of time}~\citep[e.g.,][]{Hawk85}.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{intro1}

  \caption{\textbf{Retrodiction, retrospection, and prediction.} In one's own
  life, one may draw on memory to retrospect (i.e., review or re-evaluate) the
  past or predict the future. This process is time-asymmetric, since our own
  past is (typically) observed whereas our future is not. When we make
  inferences about \textit{other} people's lives, however, we often have
  uncertainty about both their past and future, since we may have observed
  neither. We may \textit{retrodict} the unobserved past and predict the
  unobserved future of other people's lives.}

  \label{fig:intro1}
\end{figure}

Although the psychological arrow of time implies that we should be better able
to infer our past than our future, how generally does this temporal asymmetry
hold? And does the asymmetry hold only for our own experiences (due to our
memories), or is the asymmetry a general property of any real-life event
sequence? In real-world situations (and narratives) where we are
\textit{equally} ignorant of the past and future, as for \textit{other}
people's lives where we lack memories of the relevant past, are our inferences
about the past and future symmetric or asymmetric? For example, imagine that
you are meeting a stranger for the first time. At the moment of your meeting,
you lack both memories of their past and knowledge about what they might do in
the future. After your first encounter with the stranger, would you be able to
more accurately or easily form inferences about what had happened in their past
(\textit{retrodiction}) or what will happen in their future
(\textit{prediction}; Fig.~\ref{fig:intro1})? Or suppose you started watching a
movie partway through. Again, you would enter the moment of watching without
memories of prior parts of the movie. Given your observations in the present,
would your guesses about what had happened before you started watching be more
(or less) accurate than your guesses about what will happen next? In general,
when the past and future are \textit{both} unobserved, are we better at
inferring the past or the future in real-world settings? Narrative stimuli,
such as stories and movies, can provide a useful testbed for exploring several
of these questions.

Although narratives are unlikely to be confused with one's own experiences,
narratives mirror some of the structure of real-world experiences. Character
behaviors and interactions are often designed in a way that helps the audience
connect with or relate to the characters. Events in narratives also unfold in
ways that are intended to build rapport or engagement with the audience. This
might be accomplished by having events follow a believable structure that is
reminiscent of real-world experiences, or by designing the audience's
experiences in ways that communicate clear ``rules'' or ``features'' that help
to immerse the audience in the narrative's universe. The characters in a
realistic narrative can also be written to behave in ways reminiscent of
real-world people. These same aspects of narratives that authors use to drive
engagement with events and characters can lead narratives to replicate some
core aspects of real-world experiences that are typically lost or overlooked in
traditional sequence learning paradigms. Narratives can drive the audience to
build situation models~\citep{RadvCope06, ZwaaRadv98} of the narrative's
universe, or to form a theory of mind of and make predictions about the
characters~\citep{TamiThor18, KostSaxe13}. Events in narratives may unfold in a
consistent or logical way, but they also exhibit complex and meaningful
interactions across events reminiscent of real-world experiences (but not
necessarily the simple sequences traditionally used in the statistical learning
literature).

One key difference between simple artificial sequences and more naturalistic
(real or narrative) sequences is that naturalistic sequences often incorporate
other people. Despite the past and future being equally unknown to \textit{the
observer} prior to the current moment, other people, and realistic characters
in narratives, have their own psychological arrows of time. Specifically, they
have memories of their own pasts. Other people's asymmetric knowledge about
their \textit{own} pasts and futures might affect their behaviors (e.g.,
conversations). In turn, this might provide time-asymmetric clues that favor
the past~\citep[e.g., other people might talk more about their own pasts than
their futures; ][]{DemiEtal18}. If observers leverage these clues from other
people's asymmetric knowledge, then observers should also be better at
inferring the past (versus the future) of other people's lives. Alternatively,
inferences about other people's lives may be more like inferences about
artificial statistical sequences~\citep[e.g., perhaps solely relying on
statistical regularities like event schemas, scripts, or situation
models;][]{RadvCope06, ZwaaRadv98, BoweEtal79, RangRitc12, BaldEtal18}. If so,
then the accuracy of inferences about the past and the future of others' lives
should be approximately equal. We note that the aforementioned authors make no
specific claims about temporal symmetries or asymmetries. Rather, we claim that
statistical regularities might \textit{imply} symmetry (e.g., if you are on
step $n$ of an unfolding schema, this suggests you have just completed step $n
- 1$ and that you are likely to next encounter step $n + 1$).

We designed a naturalistic paradigm for exposing participants to scenarios
where the past and future were equally unobserved. We asked our participants to
watch a series of movie segments drawn from a character-driven dramatic
television show. Across the conditions and trials in the experiment,
participants made free-form text responses to either retrodict what had
happened in the previous segment, predict what would happen in the next
segment, or recall what happened in the just-watched segment. We used manual
annotations and sentence-level natural language processing models to
characterize participants' responses. To foreshadow our results, we found that
participants were overall better at retrodicting the past than predicting the
future. This appeared to be driven by two main factors. First, characters more
often referred to past events than future (e.g., planned) events, and this
influenced participants' responses. Second, associations and dependencies
between temporally adjacent events enabled participants to form estimates about
nearby events (e.g., to a just-watched scene or a past or future event
referenced in an observed conversation). We also ran a pre-registered
replication study to confirm that these findings generalized to another
television show and group of participants. Finally, we ran a meta analysis
using natural language processing to estimate the prevalence of references to
past and future events in hundreds of millions of dialogues drawn from
television shows, popular movies, novels, and written and spoken natural
conversations. Taken together, our work reveals a temporal asymmetry in how
observations of other humans’ behaviors inform us about the past versus the
future.

\section*{Results}

Participants in our main experiment ($n = 36$) watched segments from two
storylines, drawn from the CBS television show \textit{Why Women Kill}. Each
storyline comprised 11 segments (mean duration: 2.05~min; range:
0.97--3.87~min, Table~\stimDescription). We asked participants to use free-form
(typed) text responses to retrodict what had happened prior to a just-watched
segment, predict what would happen next, or recall what they had just watched
(Fig.~\ref{fig:method}, \textit{Task design}). We referred to the
to-be-retrodicted, to-be-predicted, or to-be-recalled segment as the
\textit{target segment} for each response. We systematically varied whether
participants watched the segments in forward or reverse chronological order,
and how many segments they had seen prior to making a response (see
\textit{Methods}).

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{methods}

  \caption{\textbf{Task overview.} Participants in our main experiment watched
  segments of two storylines from the television series \textit{Why Women
  Kill}. They made free-form text responses to either retrodict what had
  happened in the previous segment, predict what would happen in the next
  segment, or recall what happened in the just-watched segment. Across four
  counterbalanced sequences, we systematically varied whether participants
  watched the segments in forward or reverse chronological order, whether (or
  not) responses were cued using the main characters in the target segment, and
  which other segments participants had watched prior to making a response. For
  each segment, we collected several retrodiction, prediction, and/or recall
  responses across different experimental conditions. Experiment time is
  denoted along the vertical axis, storyline segments are indicated along the
  horizontal axis, and the colors denote experimental tasks (conditions). For
  an analogous depiction of our replication experiment's design, see
  Fig.~\MethodsReplExp.}

  \label{fig:method}
\end{figure}

We asked participants in our main experiment to generate four types of
responses after watching each video segment: uncued responses, character-cued
responses, updated responses, and recalls (Fig.~\ref{fig:method}, \textit{Data
overview}). To generate \textit{uncued} responses, we asked participants to
either retrodict (uncued retrodiction; \textit{u-R}) what happened shortly
before or predict (uncued prediction; \textit{u-P}) what happened shortly after
the just-watched segment. To generate \textit{character-cued} responses, we
asked participants to retrodict (character-cued retrodiction; \textit{c-R}) or
predict (character-cued prediction; \textit{c-P}) what came before or after the
just-watched segment, but we provided additional information to the participant
about which character(s) would be present in the target (to-be-retrodicted or
to-be-predicted) segment. We hypothesized that character-cued responses should
be more accurate than uncued responses, to the extent that participants
incorporate the character information we provided to them into their
retrodictions and predictions. To generate updated responses, we asked
participants to watch an additional segment that came just prior to or just
after the target segment, and then to update their retrodiction (\textit{c-RP})
or prediction (\textit{c-PR}) about the target segment. Results on updated
responses are not reported in this paper. Finally, we also asked participants
to \textit{recall} what happened in the just-watched segment. We labeled these
responses according to which other segments participants had watched prior to
the just-watched target. Retrodiction-matched recall (\textit{re(R)}) responses
were made during the retrodiction sequences (B1 and B2; Fig.~\ref{fig:method}),
whereas prediction-matched recall (\textit{re(P)}) responses were made during
the prediction sequences (A1 and A2; Fig.~\ref{fig:method}). Whereas
retrodiction and prediction responses reflect what participants
\textit{estimate} they would remember after watching the (inferred) target
segment, recall responses provide a benchmark for comparison by measuring what
they \textit{actually} remember about the target segment. Our replication
experiment (Fig.~\MethodsReplExp) used a similar design, but did not have
participants generate recall, re(R), or re(P) responses.

For each retrodiction and prediction, participants were asked to generate at
least one, and not more than three, responses that constituted ``the sorts of
things [the participant would] expect to have remembered if [they] had watched
the [target] segment.'' They were asked to generate multiple responses only if
those additional responses were (in their judgement) of equal likelihood to
occur. On average, participants in our main experiment generated 1.08 responses
per prompt; therefore we chose to consider only participants' first (``most
probable'' or ``most important'') responses to each prompt. We also discarded a
small number ($n = 20$) of character-cued responses that did not contain
references to all cued characters, along with one additional response due to
the participant's misunderstanding of the task instructions during that trial.
We carried out our analyses on the remaining 2084 retrodiction, prediction, and
recall responses. (Our replication experiment analyses were carried out on XXX
responses.)

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.80\textwidth]{results1}

  \caption{\textbf{Retrodiction, prediction, and recall performance by
  experimental condition in our main experiment.} \textbf{A. Methods
  schematic.} For each retrodiction, prediction, and recall response, we
  calculated the hit rate for events in the target segment, the response
  precision (see \textit{Methods}), and the response convergence across
  participants (see \textit{Methods}). \textbf{B. Target event hit rate.} Mean
  proportions of target events that were contained in participants' responses,
  for each response type, averaged across target segments. \textbf{C. Response
  precision.} Mean precisions of participants' responses, for each response
  type, averaged across target segments. The horizontal lines denote the mean
  pairwise semantic similarities (see \textit{Methods}) across recall responses
  (re(R): orange; re(P): blue). \textbf{D. Response convergence.} Mean
  (across-participant) convergence of participants' responses, for each
  response type, averaged across target segments. All panels: error bars denote
  bootstrapped 95\% confidence intervals. Asterisks indicate significance in
  the (generalized) linear mixed models: * denotes $p < 0.05$ and ** denotes $p
  < 0.01$. See Figure~\targetAsymmetries~for analogous results from our
  replication experiment.}

  \label{fig:result1}
\end{figure}

We used two general approaches to assess the quality of participants' responses
(see \textit{Methods}, Figs.~\ref{fig:result1}A). One approach entailed
manually annotating events in the video and counting the number of matched
events in participants' responses. We identified a total of 117 unique events
reflected across the 22 video segments in our main experiment (range: 3--9 per
segment; see \textit{Methods}, Table \stimDescription). We assigned one
``point'' to each of these video events. We also identified 23 additional
events in participants' responses that were either summaries of several events
or that were partial matches to the manually identified video events. We
assigned 0.5 point to each of these additional events. This point system
enabled us to compute the numbers and proportions (\textit{hit rates}) of
correctly retrodicted, predicted, and recalled events contained in each
response. Our second approach entailed using a natural language processing
model~\citep{CerEtal18} to embed annotations and responses in a 512-dimensional
feature space. This approach was designed to capture conceptual overlap between
responses that were not necessarily tied to specific events. To quantify this
conceptual overlap, we computed the similarities between the embeddings of
different sets of responses. Following \cite{HeusEtal21}, we defined the
\textit{precision} of each participants' retrodictions or predictions about a
target segment as the median cosine similarities between the embeddings of (a)
the participant's retrodiction or prediction response for the target segment
and (b) each \textit{other} participant's recalls of the same segment. In other
words, precision is designed to measure the extent to which retrodictions and
predictions captured the conceptual content that (other) participants
remembered. We also developed a related measure, which we call
\textit{convergence}, to characterize response similarities across
participants. In particular, we defined convergence as the mean cosine
similarity between the embeddings of a participant's responses to a target
segment and all other participants' responses (of the same type) to the same
segment. We analyzed the data using generalized linear mixed models, with
participant and stimulus (e.g., target segment) identities as crossed random
effects (see \textit{Methods}).

First we sought to validate a main effect of response type (i.e., uncued
responses, character-cued responses, and recalls), irrespective of the temporal
direction (retrodiction versus prediction). Across these three types of
responses, participants have access to increasing amounts of information about
the target segment. Therefore, across these response types, we hypothesized
that participants' responses should become both more accurate and more
convergent across individuals. Consistent with this hypothesis, participants'
character-cued retrodictions and predictions were associated with higher target
event hit rates than uncued retrodictions and predictions in our main
experiment (odds ratio (OR): 2.65, $Z = 4.24$, $p < 0.001$, 95\% confidence
interval (CI): 1.69 to 4.16; Fig.~\ref{fig:result1}B). These character-cued
responses were also more precise ($b = 0.13$, $t(18.1) = 9.43$, $p < 0.001$,
CI: 0.10 to 0.16; Fig.~\ref{fig:result1}C) and convergent across individuals
($b = 0.11$, $t(18.6) = 6.21$, $p < 0.001$, CI: 0.07 to 0.15;
Fig.~\ref{fig:result1}D). Relative to character-cued responses, participants'
recalls showed higher target event hit rates (OR = 21.83, $Z = 10.61$, $p <
0.001$, CI: 12.35 to 38.59) and were more convergence across individuals ($b =
0.20$, $t(19.4) = 9.10$, $p < 0.001$, CI: 0.16 to 0.25). These results are
consistent with the common-sense notion that access to more information about a
target segment yields better performance (i.e., higher hit rates, precision,
and convergence across individuals). These findings also held for our
replication experiment (Fig.~\targetAsymmetries; hit rates of character-cued
vs. uncued responses: OR: XXX, $Z = XXX$, $p = XXX$, 95\% confidence interval
(CI): XXX to XXX; precisions of character-cued vs. uncued responses: $b = XXX$,
$t(XXX) = XXX$, $p = XXX$, CI: XXX to XXX; convergence of character-cued vs.
uncued responses: $b = XXX$, $t(XXX) = XXX$, $p = XXX$, CI: XXX to XXX).

Next we carried out a series of analyses specifically aimed at characterizing
temporal direction effects--- i.e, the relative quality of retrodictions versus
predictions across different types of responses. We hoped that these analyses
might provide insights into our central question about whether inferences about
the past and future are equally accurate. Across both uncued and character-cued
responses in our main experiment (Fig.~\ref{fig:method}), retrodictions had
numerically higher hit rates than predictions (Fig.~\ref{fig:result1}B).
However, these differences were only statistically reliable for character-cued
responses (uncued responses: OR = 1.17, $Z = 0.35$, $p = 0.73$, CI: 0.47 to
2.92; character-cued responses: OR = 1.93, $Z = 2.15$, $p = 0.03$, CI: 1.06 to
3.52). We observed a similar pattern of results for the precisions of
participants' responses (Fig.~\ref{fig:result1}C). Specifically, their
responses tended to be numerically more precise for retrodictions versus
predictions, but the differences were only statistically reliable for
character-cued responses (uncued responses: $b = 0.03$, $t(20.9) = 1.09$, $p =
0.29$, CI: -0.03 to 0.10; character-cued responses: $b = 0.06$, $t(20.8) =
3.01$, $p = 0.007$, CI: 0.02 to 0.11). We also consistently observed
numerically higher convergence across participants for retrodictions versus
predictions (Fig.~\ref{fig:result1}D), but neither of these differences were
statistically reliable (uncued responses: $b = 0.03$, $t(17.9) = 0.75$, $p =
0.46$, CI: -0.05 to 0.11; character-cued responses: $b = 0.04$, $t(17.4) =
1.46$, $p = 0.16$, CI: -0.02 to 0.09). In our replication experiment
(Fig.~\targetAsymmetries), participants were numerically better at making
\textit{predictions} than retrodictions, but none of these differences were
statistically reliable (hit rate for uncued responses: OR = XXX, $Z = XXX$, $p
= XXX$, CI: XXX to XXX; hit rate for character-cued responses: OR = XXX, $Z =
XXX$, $p = XXX$, CI: XXX to XXX; precision for uncued response: $b = XXX$,
$t(XXX) = XXX$, $p = XXX$, CI: XXX to XXX; precision for character-cued
responses: $b = XXX$, $t(XXX) = XXX$, $p = XXX$, CI: XXX to XXX; convergence
for uncued responses: $b = XXX$, $t(XXX) = XXX$, $p = XXX$, CI: XXX to XXX;
convergence for character-cued responses: $b = XXX$, $t(XXX) = XXX$, $p = XXX$,
CI: XXX to XXX). Taken together, our results across our main and replication
experiment suggest that whether participants are better at retrodicting versus
predicting the immediate past or future may be somewhat stimulus specific. We
also verified that this was not solely a consequence of how participants'
memory performance might have been affected by watching different segments (or
making different responses to other segments) across conditions by comparing
recall responses in the retrodiction-matched recall (\textit{re(R)}) and
prediction-matched recall (\textit{re(P)}) conditions. Recall performance in
our main experiment was similar in both conditions (target event hit rate: OR =
1.12, $Z = 1.07$, $p = 0.29$, CI: 0.91 to 1.39; convergence: $b = 0.03$,
$t(19.3) = 1.89$, $p = 0.07$, CI: 0.00 to 0.07). (We did not collect recall
responses in our replication experiment.)

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{results2}

  \caption{\textbf{Retrodictions and predictions of temporally near and distant
  events.} \textbf{A. Illustration of annotation approach.} For each uncued
  retrodiction and prediction response in our main experiment, we calculated
  the number of (retrodicted or predicted) events as a function of temporal
  distance from the target segment, or \textit{lag}. Onscreen (explicit) events
  are tagged using integer-valued lags, whereas offscreen (implicit) events are
  tagged using half-step lags ($\pm 0.5$, $\pm 1.5$, etc.). \textbf{B. Number
  of events hit in participants' uncued retrodictions and predictions for each
  event type.} Here we separated events we identified in participants'
  responses according to whether they occurred in the target segment (lags of
  $\pm 1$), during the interval between the target segment and the just-watched
  segment (lags of $\pm 0.5$), at longer temporal distances ($|\mathrm{lag}| >
  1$), or were incorrect (unmatched with any past or future events in the
  narrative). The counts displayed in the panel are averaged across
  just-watched segments. \textbf{C. Number of events hit as a function of
  temporal distance.} Here the (across-segment) mean numbers of events hit in
  participants' uncued retrodictions (orange) and predictions (blue) are
  displayed as a function of temporal distance to the just-watched segment
  (lag). Error bars denote bootstrapped 95\% confidence intervals. Colors
  denote temporal direction (orange: past; blue: future) and distance (darker
  shading: onscreen events from segments adjacent to the target segment;
  lighter shading: offscreen events). See Figure~\hitRates~for an analogous
  presentation of results from our replication study.}

  \label{fig:result2}
\end{figure}

The above analyses were focused solely on the target segment (i.e.,
retrodiction of segment $n$ after watching segments $(n+1)... 11$, or
prediction of segment $n$ after watching segments $1 ...(n-1)$). We wondered
whether participants' responses might also contain longer-range information
about preceding or proceeding events. In order to carry out this analysis
properly, we reasoned that participants might reference past or future events
that were \textit{implied} to have occurred offscreen, but not explicitly shown
onscreen. For example, a character in location A during one scene might appear
in location B during the immediately following scene. Although it wasn't shown
onscreen, we can infer that the character traveled between locations A and B
sometime between the time intervals separating the scenes~\citep{Bord08}. In
all, we manually identified a set of 74 \textit{implicit} offscreen events in
our main experiment's stimuli that were implied to have occurred given what was
(explicitly) depicted onscreen (Fig.~\ref{fig:result2}A), plus one additional
partial event and one additional summary event. We applied the same procedure
to our replication experiment's stimuli and identified XXX implicit offscreen
events. We defined the just-watched segment as having a \textit{lag} of 0. We
assigned the target segment of a participant's retrodiction or prediction
(i.e., the immediately preceding or proceeding segment) a lag of -1 or +1,
respectively. The segment following the next was assigned a lag of 2, and so
on. We tagged offscreen events using half steps. For example, an offscreen
event that occurred after the prior segment but before the just-watched segment
would be assigned a lag of -0.5.

Because there is no ``ground truth'' number of offscreen events, we could not
compute the hit rates for offscreen events. Instead, we counted up the absolute
\textit{number} of retrodicted or predicted events as a function of lag. In
other words, given that the participant had just watched segment $i$, we asked
how many events from segment $i + lag$ they retrodicted or predicted, on
average, given that they were aiming to retrodict or predict events at lags of
$\pm 1$. We also counted the numbers of \textit{unmatched} events in
participants' responses that did not correspond to any events in the relevant
segments of the narrative. We focused specifically on \textit{uncued}
retrodictions and predictions, which we hypothesized would provide the cleanest
characterizations of participants' initial estimates of the unobserved past and
future (i.e., without potential biases introduced by additional character
information, as in the character-cued responses). For participants in our main
experiment, the numbers of uncued retrodicted and predicted target (lag = $\pm
1$) events were not reliably different (OR = 0.92, $Z = -0.15$, $p = 0.88$, CI:
0.30 to 2.84). In other words, uncued retrodictions and predictions over short
timescales did not exhibit reliable asymmetries. This ``null result'' also held
in our replication study (OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX).
However, when retrodicting, participants in both experiments mentioned events
from the distant past ($\mathrm{lag} < -1$) more often than participants
predicted events from the distant future ($\mathrm{lag} > 1$; main experiment:
OR = 9.10, $Z = 3.80$, $p < 0.001$, CI: 2.92 to 28.39; Fig.~\ref{fig:result2}B,
C; replication experiment: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX;
Fig.~\hitRates; for results from the character-cued conditions, see
Fig.~\events). Despite this asymmetry in the accuracies of participants'
long-range retrodictions versus predictions, there were no reliable differences
in the \textit{numbers} of uncued retrodicted versus predicted events (across
all lags; main experiment: OR = 1.05, $Z = 0.75$, $p = 0.45$, CI: 0.93 to 1.18;
replication experiment: OR = XXX, $Z = XXX$, $p = XXX$). Nor did we find any
reliable differences in the numbers of offscreen events immediately before or
after the just-watched segment ($lag = \pm 0.5$; main experiment: OR = 0.75, $Z
= -0.36$, $p = 0.72$, CI: 0.15 to 3.59; replication experiment: OR = XXX, $Z =
XXX$, $p = XXX$, CI: XXX to XXX). The apparent discrepancy between
participants' asymmetric accuracy but symmetric event counts was due to
participants' tendencies to reference ``unmatched'' events (i.e., events that
did not correspond to any explicit or implicit event in the story) more in
their predictions than retrodictions (main experiment: OR = 0.36, $Z = -4.53$,
$p < 0.001$, CI: 0.23 to 0.56; replication experiment: OR = XXX, $Z = XXX$, $p
= XXX$, CI: XXX to XXX). We confirmed that the retrodiction advantage held when
controlling for absolute lag (main experiment: OR = 34.31, $Z = 3.28$, $p =
0.001$, CI: 4.16 to 283.20; replication experiment: OR = XXX, $Z = XXX$, $p =
XXX$, CI: XXX to XXX), for onscreen events alone (main experiment: OR = 47.54,
$Z = 3.74$, $p < 0.001$, CI: 6.27 to 360.60; replication experiment: OR = XXX,
$Z = XXX$, $p = XXX$, CI: XXX to XXX), and marginally for offscreen events
alone (main experiment: OR = 24.76, $Z = 1.71$, $p = 0.09$, CI: 0.63 to 975.27;
replication experiment: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX). Taken
together, these analyses show that (in generating uncued responses)
participants tend to reach ``further'' into the unobserved past, and with
greater accuracy, than the unobserved future.

What might be driving participants to retrodict further and more accurately
into the unobserved past, compared with their predictions of the unobserved
future? By inspecting the video content, we noticed that characters frequently
referenced both past events and (planned or predicted) future events in their
spoken conversations. We wondered whether the characters' references might show
temporal asymmetries that might explain participants' behaviors. Across all of
the characters' conversations, and across all of the video segments from our
main experiment, we manually identified a total of 82 references to past or
future events (i.e., that occurred onscreen or offscreen before or after the
events depicted in the current segment; Figs.~\ref{fig:result3}A, \references
A, \characterRefs). Characters in our main experiment's stimulus tended to
reference the past (52 references) more than the future (30 references),
consistent with previous work~\citep{DemiEtal18}. References to the past were
also skewed to more temporally distant events compared with references to the
future (Figs.~\ref{fig:result3}B,~\references B,~\characterRefs). These
asymmetries also held for characters in the replication experiment's stimulus
(Fig.~\ref{fig:meta-analysis}). These observations indicate that the characters
in the stimulus display a ``preference'' for the past (versus future) in their
conversations. Might this asymmetry be driving the asymmetries in participants'
retrodictions versus predictions?

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.8\textwidth]{results3}

  \caption{\textbf{Characters' references drive participants' retrodiction and
  prediction performance.} \textbf{A. Illustration of annotation approach.} We
  manually annotated references to events in past or future segments in
  characters' spoken conversations. We matched each such reference with its
  corresponding storyline event (and its corresponding segment number for
  onscreen events, or half-step segment number for offscreen events). We then
  tracked the hit rate separately for referenced versus unreferenced events in
  participants' uncued retrodictions and predictions. \textbf{B. Reference rate
  as a function of lag.} Across all possible just-watched segments (lag 0), the
  bar heights denote the average proportions of events referenced in other past
  (orange, negative lags) or future (blue, positive lags) segments in our main
  experiment's stimulus. \textbf{C. Difference in hit rates between all events
  and unreferenced events.} To highlight the effect of characters' references
  to past and future events on participants' retrodictions and predictions,
  here we display the difference in across-segment mean hit rates between all
  events and unreferenced events, as a function of temporal distance (lag) to
  the just-watched segment. \textbf{D. Hit rates for unreferenced events.} The
  average response hit rates for unreferenced events are displayed as a
  function of temporal distance to the just-watched segment. Error bars denote
  bootstrapped 95\% confidence intervals. Panels B--D: colors are described in
  the Figure~\ref{fig:result2} caption. \textbf{E. Hit rates and counts of
  referenced and unreferenced events.} As a function of temporal distance to
  the just-watched segment, the sub-panels display the across-segment mean
  numbers ($x$-axes) and hit rates ($y$-axes) of referenced (red) and
  unreferenced (gray) events that participants hit (darker shading) or missed
  (lighter shading) in their uncued retrodictions (top sub-panel) and uncued
  predictions (bottom sub-panel). Intuitively, the widths of the rectangles at
  each lag denote the total number of events at each possible lag. The darker
  shading denotes the proportions of events that participants retrodicted or
  predicted, and the lighter shading denotes the proportions of events that
  participants ``missed'' in their responses. For an analogous presentation of
  results from the replication experiment, see Figure~\characterRefs.}

  \label{fig:result3}
\end{figure}

Controlling for temporal distance (lag), past and future events that story
characters referenced in their conversations were associated with higher hit
rates than unreferenced events in our main experiment (uncued retrodiction: OR
= 12.70, $Z = 10.94$, $p < 0.001$, CI: 8.06 to 20.03; uncued prediction: OR =
8.29, $Z = 6.83$, $p < 0.001$, CI: 4.52 to 15.20; Fig.~\ref{fig:result3}E).
This indicates that participants' responses are at least partially influenced
by the characters' conversations. To estimate the contributions of characters’
references on hit rates, we computed the difference in hit rates between all
events (which comprised both referenced and unreferenced events) and
unreferenced events, as a function of lag. These differences exhibited a
temporal asymmetry in favor of retrodiction (Figs.~\ref{fig:result3}C). This
indicates that the asymmetries in participants' retrodictions versus
predictions are also at least partially influenced by the characters'
conversations. However, these temporal asymmetries in participants'
retrodictions and predictions persisted even for events that characters never
referenced in their conversations (hit rates of uncued retrodicted versus
predicted unreferenced events: OR = 2.00, $Z = 2.40$, $p = 0.02$, CI: 1.14 to
3.51; Fig.~\ref{fig:result3}D). When we further separated the unreferenced
events into onscreen events and offscreen events, we found that these
asymmetries held only for the onscreen events (onscreen: OR = 2.65, $Z = 2.59$,
$p = 0.01$, CI: 1.27 to 5.54; offscreen: OR = 1.50, $Z = 0.91$, $p = 0.36$, CI:
0.63 to 3.62). We found similar patterns in our replication experiment
(Fig.~\characterRefs; hit rates of uncued retrodictions for referenced events:
OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX; uncued predictions for
referenced events: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX; hit rates of
uncued retrodictions for \textit{unreferenced} events: OR = XXX, $Z = XXX$, $p
= XXX$, CI: XXX to XXX; for predicted events: OR = XXX, $Z = XXX$, $p = XXX$,
CI: XXX to XXX). Taken together, these analyses suggest that asymmetries in the
number of references characters make to past and future events partially (but
not entirely) explain why participants tend to retrodict the past further and
more accurately than they predict the future.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.8\textwidth]{results4}

  \caption{\textbf{Reference-adjacent events are associated with higher hit
  rates (main experiment).} \textbf{A. Illustration of annotation approach.} We
  extended the annotation procedure depicted in Figure~\ref{fig:result3}A to
  also label unreferenced events that were either temporally adjacent to (i.e.,
  immediately preceding or proceeding) a referenced event (reference-adjacent
  events) or not (remaining events). \textbf{B. Adjacent reference rate for
  unreferenced events as a function of lag.} Across all possible just-watched
  segments (lag 0), the bar heights denote the average proportion of
  unreferenced events in other past (orange, negative lags) or future (blue,
  positive lags) segments that were temporally adjacent to any referenced
  event. \textbf{C. Difference in hit rates between unreferenced events and
  remaining events.} To highlight the effect of reference adjacency on
  retrodiction and prediction of unreferenced events, here we display the
  difference in across-segment mean hit rates between unreferenced events and
  remaining events, as a function of temporal distance (lag) to the
  just-watched segment. \textbf{D. Hit rates for remaining events.} The
  across-segment mean response hit rates for unreferenced events that were
  \textit{not} temporally adjacent to any referenced events are displayed as a
  function of temporal distance to the just-watched segment. Error bars denote
  bootstrapped 95\% confidence intervals. Panels B--D: colors are described in
  the Figure~\ref{fig:result2} caption. \textbf{E. Hit rates and counts of
  referenced, reference-adjacent, and remaining events.} As a function of
  temporal distance to the just-watched segment, the sub-panels display the
  numbers ($x$-axes) and proportions ($y$-axes) of referenced (red),
  reference-adjacent (purple), and remaining (gray) events that participants
  hit (darker shading) or missed (lighter shading) in their uncued
  retrodictions (top sub-panel) and uncued predictions (bottom sub-panel). For
  an analogous depiction of results from our replication experiment see
  Fig.~\refAdjacent.}

  \label{fig:result4}
\end{figure}

If characters' direct references cannot fully account for the temporal
asymmetry in retrodicting the unobserved past versus predicting the unobserved
future, what other factors might explain this phenomenon? The results above
indicate that characters’ references to specific unobserved events in the past
or future boost participants’ estimates of these events. But might characters'
references have other effects on participants' responses \textit{beyond} the
referenced events? For example, real-world experiences and events in realistic
narratives are often characterized by temporal autocorrelations (i.e., what is
``happening now'' will likely relate to what happens ``a moment from now,'' and
so on). Real-world experiences and realistic narratives are also often
structured into ``schemas'' whereby experiences unfold according to a
predictable pattern or formula that characterizes a particular situation, such
as going to a restaurant or catching a flight at the
airport~\citep{BaldEtal18}. If there are associations or temporal dependencies
between temporally nearby events in the television show participants watched,
participants might be able to pick up on these patterns in forming their
responses. This would be reflected in an inference ``boost'' for events that
were \textit{nearby in time} to events that characters referred to in their
conversations, in addition to the referenced events themselves
(Fig.~\ref{fig:result4}A).

Because characters tended to refer to past events more often than future
events, the proportions of unreferenced events that were adjacent to referenced
events should show a similar temporal asymmetry in favor of the past. We tested
this intuition by computing the proportions of unreferenced events in the
stimulus that were temporally adjacent to past or future events referenced by
the characters during a given segment. Here we defined \textit{temporally
adjacent} as any event within an absolute lag of one relative to a referenced
onscreen event, or within an absolute lag of 0.5 to a referenced offscreen
event. We also defined \textit{remaining} events as unreferenced events that
were not temporally adjacent to any referenced events. As shown in
Figure~\ref{fig:result4}B, in our main experiment we observed higher
proportions of unreferenced past than future events that were temporally
adjacent to referenced events. Further, these reference-adjacent events had
higher hit rates than remaining events after controlling for absolute lag
(uncued retrodiction: OR = 7.15, $Z = 2.40$, $p = 0.02$, CI: 1.44 to 35.58;
uncued prediction: OR = 3.11, $Z = 2.30$, $p = 0.02$, CI: 1.18 to 8.21;
Fig.~\ref{fig:result4}E). These findings also held in our replication
experiment (uncued retrodiction: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to
XXX; uncued prediction: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX;
Fig.~\refAdjacent). To estimate the contributions of reference adjacency on hit
rates, we computed the difference in hit rates between unreferenced events
(which comprised both reference-adjacent and remaining events) and remaining
events, as a function of lag. These differences exhibited a temporal asymmetry
in favor of retrodiction. This suggests that reference-adjacent events also
contribute to participants' retrodiction advantage. Remaining events did
\textit{not} exhibit a reliable temporal asymmetry (main experiment: OR = 0.75,
$Z = 0.33$, $p = 0.74$, CI: 0.14 to 4.08, Fig.~\ref{fig:result4}D; replication
experiment: OR = XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX, Fig.~\refAdjacent
D), suggesting that, after accounting for temporal adjacency, character's
references to past and future events can explain participants' retrodiction
advantage.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{results5}

  \caption{\textbf{Referenced events are associated with higher hit rates, but
  referring events are not.} \textbf{A. Illustration of annotation approach.}
  We extended the annotation procedure depicted in Figure~\ref{fig:result3}A to
  also label which events in our main experiment's stimuli \textit{contained}
  references to events in other segments. \textbf{B. Referenced versus
  referring events.} During event $i$, when a character makes a reference to
  another event ($j$), we define $i$ as the \textit{referring} event and $j$ as
  the \textit{referenced} event. \textbf{C. Referring rate as a function of
  lag.} Across all possible just-watched segments (lag 0), the bar heights
  denote the across-segment mean proportions of events containing references to
  events in other past (orange, negative lags) or future (blue, positive lags)
  segments in our main experiment's stimuli. The bar colors are described in
  the Figure~\ref{fig:result2} caption. \textbf{D. Hit rates and counts of
  referenced, referring, and other events.} As a function of temporal distance
  to the just-watched segment, the sub-panels display the numbers ($x$-axes)
  and hit rates ($y$-axes) of referenced (red), referring (green), and other
  (gray) events that participants hit (darker shading) or missed (lighter
  shading) in their uncued retrodictions (top sub-panel) and uncued predictions
  (bottom sub-panel). For a display of analogous results from our replication
  experiment see Figure~\referringReferenced.}

  \label{fig:result5}
\end{figure}

The preceding analyses show that when characters reference past or future
events, those referenced events, and other events that are temporally adjacent
to the referenced events, are more likely to be retrodicted and predicted. In
other words, referring to a past or future event in conversation leads to a
``boost'' in that event's hit rate. We wondered whether this boost was
bi-directional. In particular: when a character refers (during a
\textit{referring event}) to another event (i.e., the \textit{referenced
event}), does this boost only the referenced event's hit rate, or does the
referring event also receive a boost? We labeled each event as a ``referring
event,'' a ``referenced event,'' or a ``other event'' (i.e., not referring or
referenced; Fig.~\ref{fig:result5}A, B). We limited our analysis to references
to onscreen (explicit) events. Consistent with our analysis of the proportions
of referenced events (Fig.~\ref{fig:result3}B), the proportions of
\textit{referring} events exhibited a \textit{forward} temporal asymmetry
(Fig.~\ref{fig:result5}C). Controlling for absolute lag, we found that
referring events were associated with lower hit rates than referenced events in
our main experiment (uncued retrodiction: OR = 0.03, $Z = -4.81$, $p < 0.001$,
CI: 0.01 to 0.11; uncued prediction: OR = 0.04, $Z = -5.84$, $p < 0.001$, CI:
0.01 to 0.12; Fig.~\ref{fig:result5}D) and had no reliable differences in hit
rates compared with other events (uncued retrodiction: OR = 0.37, $Z = -1.46$,
$p = 0.15$, CI: 0.10 to 1.41; uncued prediction: OR = 2.16, $Z = 1.68$, $p =
0.09$, CI: 0.88 to 5.30). We also observed this phenomenon in our replication
experiment (referenced events, uncued retrodiction: OR = XXX, $Z = XXX$, $p =
XXX$, CI: XXX to XXX; referenced events, uncued prediction: OR = XXX, $Z =
XXX$, $p = XXX$, CI: XXX to XXX; other events, uncued retrodiction: OR = XXX,
$Z = XXX$, $p = XXX$, CI: XXX to XXX; other events, uncued prediction: OR =
XXX, $Z = XXX$, $p = XXX$, CI: XXX to XXX; Fig.~\referringReferenced). Taken
together, this indicates that only referenced events received a hit rate boost
(relative to other events), suggesting that the retrodictive and predictive
benefits of references are directed (i.e., asymmetric).

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{meta-analysis}

\caption{\textbf{Meta analysis.} We used natural language processing to
automatically identify references to past or future events across a variety of
sources. \textbf{A. Numbers of past and future events in \textit{The Chair},
Season 1, Episodes 1--6.} The bar heights indicate the raw numbers of manually
identified (lighter shading) and automatically identified (darker shading) past
and future events from each episode (color). We used Episode 1 from this series
as the stimulus in our replication experiment. \textbf{B. Proportions of past
and future events in \textit{The Chair}, Season 1, Episodes 1--6.} The Panel is
in the same format as Panel A, but here the bar heights have been divided by
the total numbers of past and future events (per episode). \textbf{C.
Proportions of past and future events in movies, television shows, and natural
conversations.} As in Panel B, the bar heights denote the proportions of past
and future events detected in each dataset (color). The datasets are described
in Table~\metaAnalysisDatasets. Error bars denote bootstrap-estimated 95\% confidence
intervals.}

  \label{fig:meta-analysis}
\end{figure}

The above analyses show that characters in the television shows we used as
stimuli in our main experiment and replication experiment refer more often to
the past than to the future. This appears to bias participants' inferences
about the past and future. But how universal is this pattern? For example, were
the television shows we happened to select for our experiment representative of
television shows more generally? Or perhaps narratives created for
entertainment purposes tends to have a bias towards the past in order to keep
the story engaging and unpredictable. To better understand temporal biases in
conversations, we carried out a meta analysis using extracted conversation data
from several large datasets, comprising over 17 million documents. The data
comprised transcripts from television shows and popular films, novels, and
spoken and written utterances from natural conversations. A summary of the data
we analyzed may be found in Table~\metaAnalysisDatasets. As summarized in
Figure~\ref{fig:meta-analysis}, we used natural language processing to identify
references to past or future events in each conversation (also see \textit{Meta
analysis of conversation data}).

To validate our basic approach, we compared the numbers
(Fig.~\ref{fig:meta-analysis}A) and proportions (Fig.~\ref{fig:meta-analysis}B)
of automatically and manually identified references to past and future events,
across six episodes of the television show \textit{The Chair}. (The first
episode was used as the stimulus in our replication study.) In general, our
automated tagging procedure tended to overcount the numbers of references. From
manually ``spot checking'' hundreds of example tags, we noticed that our
automatted tagging procedure often counts the ``same'' references multiple
times. Specifically, the manually generated tags sought to identify references
to specific events that occurred or were implied to occur in other parts of the
narrative. In contrast, as a heuristic, we designed the automatic tagging
procedure to identify uses of the past or future \textit{tense} as a proxy for
references to past or future \textit{events}. Individual conversations often
contains multiple references to a given (past or future) event. Whereas the
manually generated tags counted these as ``single'' references, our automated
tagging procedure has no means of differentiating between several references to
the same event versus the same number of references to different events. This
leads the automated tagging procedure to overestimate the numbers of distinct
events being referenced. Nevertheless, this discrepancy did not appear to bias
the balance of the overall \textit{proportions} of past or future references.

In all, across all of the datasets we examined in our meta analysis, we
identified a total of 36,008,500 references to past or future events. A total
of 19,464,741 (54.06\%) of these were references to past events, and the
remaining 16,543,759 (45.94\%) were references to future events. We also
computed the average proportions of references to past and future events across
documents within each individual dataset. Across the 12 datasets we examined
(Fig.~\ref{fig:meta-analysis}, Tab.~\metaAnalysisDatasets), there were
significantly more references to the past than the to the future (mean $\pm$
standard deviation proportion of references to past events: $58.99\% \pm
7.28\%$; $t(11) = 4.28, p = 0.0013$). This bias towards the past also held for
each dataset individually ($t$s $\geq 5.14$, $p$s $< 0.01$) except for one
dataset, ``Persuasion for Good,'' which comprised natural conversations between
pairs of Amazon Mechanical Turk workers wherein one participant tried to
convince the other participant to donate to a charity in the future. In that
dataset, references to the future were significantly more common than
references to the past ($t(11438) = -22.65, p < 0.001$). This latter example
provided a nice sanity check for verifying that our general approach was not
itself biased in favor of the past, e.g., even in conversations that were
actually biased towards the future. Taken together, the results from our meta
analysis indicate that people tend to refer to the past more than they refer to
the future, across a wide variety of situations (including in both fictional
and real conversations). Although (as in the Persuasion for Good dataset) there
may be specific exceptions to this bias, it seems that a bias in favor of the
past is a common element of many (and perhaps even \textit{most}) human
conversations.

\section*{Discussion}

We asked participants in our main experiment to watch sequences of movie
segments from a character-driven television drama and then either retrodict
what had happened prior to a just-watched segment, predict what would happen
next, or recall what they had just watched. We found that participants tended
to more accurately and more readily retrodict the unobserved past than predict
the unobserved future. We traced this temporal asymmetry to (a) characters'
tendencies to refer to past events more than future events in their ongoing
conversations, and (b) associations between temporally proximal events
(Fig.~\ref{fig:discussion}). Essentially, associations between temporally
proximal events serve to enhance asymmetries in inferences driven by
conversational references (light orange and blue bars in
Fig.~\ref{fig:discussion}). Our findings show that other peoples' psychological
arrows of time can affect external observers' inferences about the unobserved
past and future. We confirmed our main behavioral findings in a pre-registered
replication study. We also carried out a meta analysis of tens of millions of
utterances from television shows, movies, novels, and natural spoken and
written conversations. We found that the tendency to refer more often to
the past than the future appears to be a widespread characteristic of human
conversation.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{discussion}

  \caption{\textbf{How much information about the past and future can be
  inferred by observing the present?} By definition, let us say that the
  present moment (lag 0) contains all information about itself (dark gray).
  Given learned statistical regularities, one might extrapolate from the
  present moment into the past or future (light gray). As illustrated in this
  schematic, the information contained in the present about other moments in
  time falls off with absolute lag. This falloff is approximately
  time-symmetric. References in the present to past events (dark orange) or
  future events (dark blue) provide additional information about those
  referenced moments in time, beyond what could be inferred solely from
  statistical regularities. This additional information about those referenced
  moments can also be extrapolated to other moments that are temporally nearby
  to \textit{them} (light orange and blue). The data in this schematic are
  hypothetical.}

  \label{fig:discussion}
\end{figure}

When people communicate through language or other observable behaviors, they
can transmit their knowledge and memories to others~\citep{HirsEcht12,
MahrCsib18, Dess07, ZadbEtal17}. A consequence of this sharing across people is
that biases or limitations in one person's knowledge and memories may also be
transmitted to external observers. Although people \textit{can} communicate
their intentions and future plans (i.e., information about their future),
because people know \textit{more} about their pasts than their futures, the
knowledge transmitted to observers is inherently biased in favor of the
past~\citep[Fig.~\ref{fig:discussion}; ][]{DemiEtal18}. Since observers
leverage communicated knowledge to reconstruct the unobserved past and future,
this explains why observers' inferences about observed people's lives also
favor the past.

People's knowledge asymmetries are not always directly observable. For example,
in a conversation where someone talks exclusively about their future plans, a
passive observer might gain more insight into the speaker's unobserved future
than their unobserved past. However, because the speaker is also guided by
their own psychological arrow of time, the ``upper limit'' of knowledge about
their past is still higher than that of their future. Therefore, after
accounting for knowledge that \textit{could} be revealed through active
participation in the conversation, the seemingly future-biased conversation
masks an underlying knowledge asymmetry in favor of the past. This hypothesized
``unmasking'' effect of interaction implies that the influence of other
people's psychological arrows of time should be more robust when the receiver
is an active participant in the conversation. Other social dimensions, such as
trust, motivation or level of engagement, personal goals, and beliefs, might
serve to modulate the effective ``gain'' of the communication channel-- i.e.,
how much the speaker's knowledge influences the observer's knowledge. Some
recent work~\citep[e.g.,][]{TamiMitc13, MeyeEtal19b} also suggests that people
might gain insights into other people using ``mental simulations'' of how they
might respond in particular situations (e.g., in the future), or of which sorts
of prior experiences might have led someone to behave a particular way in the
present.

In typical statistical sequences used in laboratory studies, there is no
temporal asymmetry, either theoretically~\citep{Cove94, BialEtal01,
ElliEtal09}, or empirically~\citep{JonePash07}. What makes narratives and
real-world event sequences time-asymmetric? Of course there are many
superficial differences between simple laboratory-manufactured sequences and
real-world experiences. As one example, real-world experiences often involve
other people who have their own memories and goals. At a deeper level, however,
are our subjective experiences essentially more complicated versions of
laboratory-manufactured sequences? Or are there fundamental differences? One
possibility is that real-life event sequences are not stationary~\citep[i.e.,
not in equilibrium; ][]{Cove94}. For example, real-life events might start from
a special initial condition~\citep{Albe00, Feyn65, Cove94} and proceed through
a series of transitions from more-ordered to less-ordered states, thus
exhibiting an arrow time. When we retrodict, it is possible that we only
consider possible past events that are compatible with the highly-ordered
special initial state~\citep{Carr10, Carr16}. For example, when we see a broken
egg we might infer that the egg had been intact at some point in the past. But
it would be difficult to guess at what states or forms the broken egg might
take in the future~\citep{Carr10, Carr16}. In other words, the procession from
order to disorder might result in better retrodiction performance compared with
that of (implicitly less-restricted) prediction tasks. The special initial
state might also explain why we remember the past, but not the future. Some
recent work suggests that the psychological arrow of time might be explained by
a related concept in the statistical physics literature, termed the
``thermodynamic'' arrow of time~\citep{MlodBrun14, Rove22}. However, the
relation between the thermodynamic and psychological arrows of time is still
under debate~\citep{Golo21, HemmShen19}.

Beyond forming inferences about unobserved past and future events, our work
also relates to prior studies of how people perceive time~\citep{BlocGrub14,
Howa18, Eagl08, IvrySchl08, Wear16}, and how we ``move'' through time in our
memories of our past experiences~\citep{Mann21a, MannEtal11, HowaEtal12,
MannEtal07, ShanHowa12, Kaha96, PolyKaha08, SchaTulv94} or in our imagined
(past or future) experiences~\citep{Scha12, JossTone20, SchaEtal98,
MomeHowa18}. For example, a well-studied phenomenon in the episodic memory
literature concerns how remembering a given event cues our memories of other
events that we experienced nearby in time~\citep[i.e., the \textit{contiguity
effect};][]{Kaha96}. Across a large number of studies there appears to be a
nearly universal tendency for people to move \textit{forwards} in time in their
memories, whereby recalling an ``event'' (e.g., a word on a previously studied
list) is about twice as likely to be followed by recalling the event that
immediately followed as compared with the event immediately preceding the
just-recalled event~\citep{HealKaha14a}. Superficially our current study
appears to report the \textit{opposite} pattern, whereby participants display a
\textit{backwards} temporal bias. However, the two sets of findings may be
reconciled when one considers the frame of reference~\citep[and current mental
context; e.g.,][]{HowaKaha02a} of the participant at the moment they make their
response. In our study, participants observe an event in the present, and they
make guesses about what happened in the unobserved past or future, relative to
the just-observed event. (Our findings imply that participants are more facile
at moving backwards in time than forwards in time, relative to ``now.'') In
contrast, the classic contiguity effect in episodic memory studies refers to
how people move through time relative to a just \textit{remembered} event. The
forward asymmetry in the contiguity effect follows from the notion that the
moment of remembering has greater contextual overlap with events \textit{after}
the remembered event from the past than events that happened before
it~\citep[for review also see ][]{MannEtal15, Mann20}.

In our study, we explicitly designed participants' experiences such that both
the past and future were unobserved. How representative is this scenario of
everyday life? For example, we might try to speculate about the unobserved
future when making plans or goals, but when might we encounter situations where
the past is unobserved but still useful for us to speculate about? Real-life
events have long-range dependencies. In general, because the future depends on
what happened in the past, discovering or estimating information about the
unobserved past can help us form predictions about the future. We illustrate
this point in Figure~\ref{fig:discussion} by showing that the additional
information contributed by a referenced past event can also extend into the
future (light orange bars at lags $> 0$). This might explain why humans devote
substantial effort and resources to attempting to figure out what happened in
the unobserved past: history, anthropology, geology, detective and forensic
science, and other related fields are each primarily focused on understanding,
retrodicting, or reconstructing unobserved past events.

\section*{Methods}
\subsection*{Participants}

\paragraph*{Main experiment.} A total of 36 participants (25 female, mean age
21.47 years, range 19--50 years) were recruited from the Dartmouth College
community for our main experiment. All participants had self-reported normal or
corrected-to-normal vision, hearing, and memory, and had not watched any
episodes of \textit{Why Women Kill} before the experiment. Participants gave
written consent to enroll in the study under a protocol approved by the
Committee for the Protection of Human Subjects at Dartmouth College.
Participants received course credit or monetary compensation for their time.
Two participants completed only the first half of the study and one
participant’s data from the second half of their testing session was lost due
to a technical error. All available data were used in the analyses.

\paragraph*{Replication experiment.} A total of XXX participants (XXX female,
mean age XXX years, range XXX--XXX years) were recruited from the Dartmouth
College community for our pre-registered replication experiment. All
participants had self-reported normal or corrected-to-normal vision, hearing,
and memory, and had not watched any episodes of \textit{The Chair} before the
experiment. Participants gave written consent to enroll in the study under a
protocol approved by the Committee for the Protection of Human Subjects at
Dartmouth College. Participants received monetary compensation for their time.
All available data were used in the analyses.

\subsection*{Stimuli}

\paragraph{Main experiment.}
The stimuli used in our main experiment were segments of the CBS television series
\textit{Why Women Kill} Season 1. The TV series contained three distinct
storylines depicting three women’s marital relationships. The three storylines,
which took place in the 1960s, 1980s, and 2019, were shown in an interleaved
fashion in the original episodes. The first 11 segments from the 1960s and
1980s storylines, across the first and second episodes, were used in our study.
Segments were divided based on major scene cuts, which primarily corresponded
to storyline shifts in the original episodes. The mean length of the segments
was 2.05 min (range 0.97--3.87 min). We chose this TV series based on its
strictly linear storytelling (within each storyline) and its realistic settings
where most events depicted everyday life. The plots were focused on the main
characters (Beth in storyline 1 and Simone in storyline 2), who were present in
all the segments in the corresponding storylines.

\paragraph{Replication experiment.} The stimuli used in our replication
experiment were segments of the first episode of the Netflix television show
\textit{The Chair}, Season 1. \textbf{JRM NOTE: Describe the show, like you did
for Why Women Kill.} The mean length of the segments was XXX min (range
XXX--XXX min). As for the stimulus we used in our main experiment, we chose
this stimulus for our replication experiment for its linear storytelling
(again, within each storyline) and its realistic depictions of everyday events.
\textit{JRM NOTE: The plots were focused on... (fill in something analogous to
what you wrote for the main experiment stimulus...)}

\subsection*{Task design and procedure}

\paragraph{Main experiment.}
Our experimental paradigm was divided across two testing sessions. In each
session, participants performed a sequence of tasks on segments from one
storyline (Fig.~\ref{fig:method}). For each storyline, there were four
different task sequences: two forward chronological order sequences and two
backward chronological order sequences. Participants completed one task
sequence in forward chronological order for one storyline, and one in backward
chronological order for the other storyline. The order of the two sessions
(forward chronological order sequence first or backward chronological order
sequence first), and the pairing of task sequences with storylines, were
counterbalanced across participants.

Tasks in each sequence alternated between watching, recall, and retrodiction or
prediction, with the specific order of tasks differing across the four
sequences. For example, in sequence A1, participants first watched segment 1,
followed by an immediate recall of segment 1. Then they predicted what would
happen in segment 2 (first uncued and then character-cued). Participants then
watched segment 3 and recalled segment 3. After that, participants guessed what
happened in segment 2 again, which we termed ``updated prediction''. Then they
watched segment 2, recalled segment 2, and so on as depicted in
Figure~\ref{fig:method}. This procedure was repeated to cover all possible
segments. We also note several edge cases at the start and end of the narrative
sequences. Since no segments precede the first segment, participants could
never make ``prediction'' responses with the first segment as their target. For
analogous reasons, participants never made ``retrodiction'' responses with the
last segment as their target. Another edge case occurred in task sequences B2
and A2 (Fig.~\ref{fig:method}). In the A1 and A2 sequences, participants
experience the narrative in the original (forward) order, predicting one
segment ahead along the way. In the B1 and B2 sequences, participants
experience the narrative in the reverse order, retrodicting one segment ahead
along the way. However, because A2 and B2 are offset from A1 and B2 by one
segment, the initial A2 responses are \textit{retro}dictions, and the initial
B2 responses are \textit{pre}dictions (i.e., they conflict with the temporal
directions of the remaining responses in those conditions). We therefore
excluded from our analysis those initial retrodiction responses from the A2
condition, and the initial prediction responses from the B2 condition.

Before watching each segment, participants were given the following task
instructions. After watching the video, participants were instructed to type
their responses (retrodiction, prediction, or recall) in 1--4 sentences.
Participants were also asked to specify the characters' names in their
responses, i.e., avoiding use of characters' pronouns. For the recall task, the
names of the characters in the recall segment were displayed, and participants
were asked to summarize the major plot points in the present tense. For the
retrodiction and prediction tasks, participants were instructed to retrodict or
predict the major plot points of the segment (also in the present tense), as
though they had watched the segment and were writing a plot synopsis. They were
also instructed to avoid speculation words (e.g., ``I \textit{think} Beth
will...''). For the uncued retrodiction and prediction tasks, participants made
retrodictions or predictions without any cues provided, so they had to guess
which of the characters would be present in the segment. For character-cued
retrodictions and predictions, the characters in the target segment were
revealed on the screen, alongside participants’ previous responses.
Participants were instructed to include or incorporate those characters into
their character-cued responses, if their previous responses did not contain all
the characters provided. They were also told that the characters were not
necessarily listed in their order of appearance in the segment, and that only
the main characters would be given. Also, the characters given did not
necessarily interact with each other in that segment, and they could appear in
successive events in that segment. If participants’ previous responses included
all the characters given, then they could directly proceed to the next task
without updating their responses. For all of the prediction and retrodiction
tasks, participants were instructed to provide at least one response, but they
were given the opportunity enter up to three responses if they felt that
multiple possibilities were more or less equally likely. Each response
(including recall) was followed by a confidence rating on a 1--5 point scale.
However, these confidence data were not analyzed in the present study.

Before their first testing session, participants were given a practice session,
where they watched the first segment of storyline 3 followed by a recall trial,
an uncued prediction trial, and a character-cued prediction trial.
Participants' responses were checked by the experimenter to ensure compliance
with the instructions. To provide participants with sufficient background
information about the storyline (especially for the backward chronological
sequences), at the beginning of each session, participants were shown the time,
location, and the main characters (with pictures) of the storyline. The first
session was approximately 1.5~h long and the second session was approximately
1~h long. We allowed participants, at their own discretion and convenience, to
sign up for two consecutive testing time-slots (i.e., with their testing
sessions occurring in immediate succession), or for testing sessions on two
different days. The mean inter-session interval was 0.73 days (range: 0--4
days). The experiment was conducted in a sound- and light-attenuated testing
room. Videos were displayed using a 27-inch iMac desktop computer (resolution:
$5120 \times 2880$) and sound was presented using the iMac’s built-in speakers.
The experiment was implemented using jsPsych~\citep{deLe15} and
JATOS~\citep{LangEtal15}.

\paragraph{Replication experiment.} \textbf{JRM NOTE: briefly describe
replication experiment methods, referring to Fig.~\MethodsReplExp. Since the
methods will have been similar for the replication study, just highlight the
differences rather than re-describing everything.}

\subsection*{Video annotation}

\paragraph{Main experiment.}
Events in the first 11 segments of the two storylines were identified by the
first author (X.X.), corresponding to major plot points (total: 117; mean: 5.32
per segment; range 3--9). Additionally, 74 offscreen events were identified. Of
these 74 offscreen events, 43 events were identified from references in
conversations during onscreen events. Another 16 events were identified based
on characters’ implied movements and travels. For example, if in segment 1
character A was in place A and in segment 2 she was in place B, then the
transit from place A to B for character A would be identified as an offscreen
event. The remaining 15 offscreen events were identified based on logical
inferences. For example, if a photograph was shown in an onscreen event (but
not the act of the photograph being taken), then the action that someone took
the photograph would be identified as an offscreen event. Offscreen events
always occurred between two contiguous segments, or before the first segment.
The purpose of identifying offscreen events was to match participants’
responses to video events; thus our identification of these offscreen events
was not intended to be exhaustive.

\paragraph{Replication experiment.} Events in the first XXX segments of the two
storylines were identified by the first author (X.X.), corresponding to major
plot points (total: XXX; mean: XXX per segment; range XXX--XXX). Additionally,
XXX offscreen events were identified. Of these XXX offscreen events, XXX events
were identified from references in conversations during onscreen events.
Another XXX events were identified based on characters’ implied movements and
travels. The remaining XXX offscreen events were identified based on logical
inferences.

\subsection*{Response analyses}

Participants' retrodiction, prediction, and recall responses were minimally
processed to correct obvious typos (e.g., in characters’ names) and remove
speculation descriptions (e.g., ``I predict that...''). All responses were
manually coded and matched to events from the video annotations. Retrodiction
and prediction responses were coded by two coders (main experiment: X.X. and
Z.Z.; replication experiment: X.X. and X.Z.). Recall responses were coded by
one coder (X.X.). While most responses were clearly identifiable as either
matching specific storyline events or as not matching any storyline events,
several ambiguous cases arose. First, some responses combined or summarized
over several (distinct) storyline events. Second, some responses lacked any
specific detail (e.g., ``character A and B talk'' without describing the
specific topic(s) of conversation or providing other relevant details). Based
on participants' responses, in addition to the original 117 onscreen events and
74 offscreen events in the main experiment's stimulus, we added 25 new events
(23 onscreen, 2 offscreen) that either summarized across several events or
partially matched the annotated events. In our replication study we used the
same procedure to add XXX new events (XXX onscreen, XXX offscreen). Whereas the
original events were each assigned a value of one point, we assigned these
additional events a half point. This point system enabled us to directly match
events in participants' responses to the annotated events. In our analyses of
retrodictions, predictions, and recalls, we added up the number of points
earned for each response to estimate participants' event hit rates.

We coded only the first retrodiction or prediction response in each trial. For
these responses, we also only considered storyline events that were in the same
temporal direction as the target segment. For example, if a participant was
asked to retrodict what happened in segment $n$, only events from segments
$1...n$ were considered in our analysis. When coding recall responses, we
considered only events from the target segment.

An additional ambiguous case arose in one main experiment participant's
responses pertaining to segment 12, storyline 2, whereby the participant
correctly identified an onscreen event that had not been included in our
original annotations. To account for this participant's response, we
retroactively added that event to our annotations of that segment. We also
identified and counted unmatched events in participants' responses (i.e.,
events that did not match any annotated events). Cases where the two coders'
independent scoring disagreed were resolved through discussions between the two
coders.

To estimate the semantic similarities between pairs of responses, we first
transformed each response into a 512-dimensional vector (embedding) using the
Universal Sentence Encoder~\citep[Transformer USE, ][]{CerEtal18}. We defined
\textit{similarity} as the cosine of the angle formed by the responses'
vectors. Following~\cite{HeusEtal21}, we defined the \textit{precision} of
participants' responses as the median similarity between that response's vector
and the embedding vectors for all other participants' recalls of the target
segment. We defined the \textit{convergence} of a given response as the mean
similarity between that response's vector and all other participants' responses
to the corresponding segment, in the same condition. To compute these median or
mean similarities we first applied the Fisher $z$-transformation to the
similarity values, then took the median or mean of the $z$-transformed
similarities, and finally applied the inverse $z$-transformation to obtain the
precision or convergence score.

To test the validity and reliability of the USE embeddings, we performed a
classification analysis of recall responses using a leave-one-out approach. For
each recall response, we calculated its semantic similarity with all other
recall responses for the same storyline. We took the segment with the highest
median semantic similarity (to the recall response) as the ``predicted''
segment. Across all responses, the predicted segments matched the true recalled
segments' labels 98.6\% of the time (1088 out of 1103 predictions; chance
level: 9\%). We note that this validation analysis could only be carried out
with data from our main experiment, since we did not collect recall responses
in our replication experiment.

\subsection*{Reference coding}

Two coders (main experiment: X.X. and Z.Z.; replication experiment: X.X. and
X.Z.) identified character dialogues in the narrative that referred to past
events or future (onscreen or offscreen) events. Only references to events that
occurred in a different segment were included in this tagging procedure. For
each reference, the source (referring) segment and the referred event number
were recorded. A total of 82 references were identified in the main experiment
stimulus, and XXX were identified in the replication experiment stimulus. Of
these references in the main experiment, 30 referred to onscreen events and 52
referred to offscreen events. In the replication experiment, XXX referred to
onscreen events and XXX referred to offscreen events. For these referenced
events, their corresponding summary events or partial events were also labelled
as referenced. In instances where the coders disagreed about a given tag,
disagreements were resolved through discussions between the two coders. In our
analyses, each storyline event was coded according to whether or not it had
been referenced in the segment(s) that the participant had viewed thus far in
the experiment.

In principle, a given event could receive multiple labels. For example, during
event $A$, a character might speak about another event, $B$, during which a
reference to a third event ($C$) was made. In this scenario, event $B$ could be
both a ``referring event'' ($B \rightarrow C$) \textit{and} a referenced event
($A \rightarrow B$). In practice, however, this scenario was quite rare,
accounting for only one out of a total of 30 onscreen events in our main
experiment and XXX events in our replication experiment.

\subsection*{Statistical analysis}

We used (generalized) linear mixed models to analyze the hit rates and numbers
of events retrodicted, predicted, and recalled, as well as the precisions and
convergences of participants' responses. Our models were implemented in
\texttt{R} using the \texttt{afex} package. We carried out comparisons or
contrasts, and extracted $p$-values, using the \texttt{emmeans} package.
Participants and stimuli (e.g., segment identity) were modeled as crossed
random effects (as specified below). Random effects were selected as the
maximal structure that allowed model convergence. All of our statistical tests
were two-sided.

For our tests of the target event hit rates across four \texttt{levels}
(uncued, character-cued, updated, and recall; Fig.~\ref{fig:result1}B), we fit
a generalized linear mixed model with a binomial link function:
\begin{lstlisting}[language=R]
  cbind(thp, ttp - thp) ~ direction * level * seg_cnt * storyline +
(direction * level | target) +
(direction * level * seg_cnt | subject)
\end{lstlisting}
where for analyses of our main experiment \texttt{thp} was the number of points hit for the target segment,
\texttt{ttp} was the total number of points for the target segment (from its
annotations), \texttt{direction} was either retrodiction or prediction,
\texttt{level} had four levels (uncued, character-cued, updated, and recall),
\texttt{seg\_cnt} represented the number of segments in the storyline that had
been watched (1--10, centered), \texttt{storyline} had two levels (1 or 2), and
\texttt{target} had 22 levels according to the identity of the target segment.
For our analyses of our replication experiment, \texttt{level} had two levels (uncued and
character-cued), the \texttt{storyline} parameter was omitted since there was 
only a single storyline, and \texttt{target} had XXX levels according to the identity of the
target segment.

For our tests of precision and convergence (Fig.~\ref{fig:result1}C, D), we fit
linear mixed models using the same formula. To test the effect of
\texttt{direction} (retrodiction or prediction) on target event hit rates,
precision, and convergence, we fit a (generalized) linear mixed model
separately for each of the three levels (uncued, character-cued, and recall).

For our tests comparing the numbers of hits for different types of events
(Fig.~\ref{fig:result2}B), we fit generalized linear mixed models using the
same formula, but with a Poisson link function. For these models, we manually
doubled the point counts to ensure that half points were mapped onto integers,
ensuring compatibility with the Poisson link function.

For our analyses of the numbers of events hit, controlling for lag
(Fig.~\ref{fig:result2}C), we fit a generalized linear mixed model with a
Poisson link function:
\begin{lstlisting}[language=R]
  hp_lag ~ direction * full_stp * lag * storyline +
  (direction | base_seg) + (1 | base_seg_pair) +
  (direction * full_stp * lag * storyline | subject)
  \end{lstlisting}
where \texttt{hp\_lag} is the number of ``points'' earned (for each lag) in
each trial (we manually doubled the point counts to ensure that half points
were mapped onto integers, for compatibility with the Poisson link function),
\texttt{full\_stp} denoted whether the given events (of the given lag) were
onscreen (i.e., full step) or offscreen (i.e., half step), \texttt{lag} denotes
the (centered) absolute lag, \texttt{base\_seg} denotes the identity of the
just-watched segment (main experiment: 22 levels; replication experiment: XXX levels), and \texttt{base\_seg\_pair} denotes the
pairing of the just-watched segment and the segment at each lag
(main experiment: 440 levels; replication experiment: XXX levels).

For our analyses of the proportions of events hit for referenced versus
unreferenced events (Fig.~\ref{fig:result3}D, E), we fit a generalized linear
model with a binomial link function:
\begin{lstlisting}[language=R]
  cbind(hp_lag, tp_lag - hp_lag) ~ direction * reference * full_stp +
  lag + (direction | base_seg) +
  (1 | base_seg_pair) +
  (direction * reference * full_stp + lag | subject)
\end{lstlisting}
where \texttt{hp\_lag} denotes the number of earned hit points for each
reference type (referenced or unreferenced) at each lag, \texttt{tp\_lag}
denotes the total number of possible hit points for each reference type at each
lag, and the other variables adhered to the same notation used in the above
formulas.

For our tests of the proportions of events hit for all three reference types
(referenced, reference-adjacent, and remaining: Fig.~\ref{fig:result4}D, E; or
referenced, referring, and other: Fig.~\ref{fig:result5}D), we fit a
generalized linear mixed model using the same formula as above, but with three
(rather than two) \texttt{reference} levels.

Several of our analyses entailed comparing the relative hit rates or
probabilities of two different conditions or outcomes. We used the
\texttt{emmeans} package to compute the odds ratios given the generalized
linear mixed models we fit for the given analysis. These odds ratios reflect
the chances (``odds'') of a particular outcome (e.g., making a response about a
particular event) given a scenario (e.g., the event occurred \textit{prior} to
the just-watched segment) compared with the chances of the outcome occurring in
the alternative scenario (e.g., the event occurred \textit{after} the
just-watched segment).

\subsection*{Meta analysis}

At a high level, the goal of our meta analysis was to predict in-text
references to past and future events. Manually identifying these references is
labor and time intensive, so it is impractical to scale up manual tagging to
millions of documents. Instead, we defined a set of heuristics for
\textit{predicting} when text is referring to real or hypothetical past or
future events. Our approach comprises four main steps.

First, we use the \texttt{nltk} package~\citep{BirdEtal09} to segment each
document into individual sentences. Each sentence is processed independently of
the others. Second, we handle contractions using the \texttt{contractions} package
(e.g., ``we'll'' is split into ``we will,'' and so on).  Third, we define two
sets of ``keywords'' (words and phrases) that tend to be indicative of referring
to the past (Tab.~\pastKeys) or future (Tab.~\futureKeys). We used
ChatGPT~\citep{ChatGPT} to generate each list, with exactly 50 templates per list,
using the following prompt:

\begin{quote}
\texttt{I'm designing a heuristic algorithm for identifying references (in text) to
past and future events. Part of the algorithm will involve looking for specific
keywords or phrases that suggest that the text is referring to something that
happened (or will happen) in the past and/or future. Could you help me generate
a list of 50 keywords or phrases to include in each list (one list for
identifying references to the past and a second list for identifying references
to the future)? I'd like to be able to paste the lists you generate into two
plain text documents with one row per keyword or phrase, and no other content.
Please output the lists as a "code" block (enclosed by '''...''').}
\end{quote}
Fourth, we use part-of-speech tagging (again, using the \texttt{nltk} package) to look for
verbs or verb phrases that are in past or future tenses.  After the words were
tagged with their predicted parts of speech, we use regular expressions
(applied to the sequences of tags) to label each verb or verb phrase with a
human readable verb form (e.g., ``future perfect continuous passive,''
``conditional perfect continuous passive,'' and so on).  The regular expressions
we used to generate these labels are shown in Table~\regExpTable, and the part of
speech tags are defined in Table~\posTags.

We treated each keyword match (of past or future keywords) as a single
``reference'' (to a past or future event, respectively), and if any past or
future verb forms were detected we treat those as (up to) one additional
reference. We then tallied up the numbers of past and/or future references
across sentences within the given document. The meta analysis results reported
in Figure~\ref{fig:meta-analysis}C display the average numbers of references
aggregated across all documents within each dataset we analyzed (described in
Tab.~\metaAnalysisDatasets).

\section*{Code and data availability}

All of the code and data generated for the current manuscript are available
online at:

https://github.com/ContextLab/prediction-retrodiction-paper

% \bibliographystyle{apa}
\bibliography{cdl}

\section*{Acknowledgements}

We thank Luke Chang, Yi Fang, Paxton Fitzpatrick, Caroline Lee, Meghan Meyer,
Lucy Owen, and Kirsten Ziman for feedback and scientific discussions. Our work
was supported in part by NSF CAREER Award Number 2145172 to J.R.M. The content
is solely the responsibility of the authors and does not necessarily represent
the official views of our supporting organizations. The funders had no role in
study design, data collection and analysis, decision to publish, or preparation
of the manuscript.

\section*{Author contributions}

Conceptualization: X.X. and J.R.M.; Methodology: X.X. and J.R.M.; Software:
X.X.; Analysis: X.X., Z.Z., X.Z., and J.R.M.; Writing, Reviewing, and Editing:
X.X., Z.Z., X.Z., and J.R.M.; Supervision: J.R.M.

\section*{Competing interests}

The authors declare no competing interests.

\end{document}